---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# symengine

[![Travis-CI Build Status](https://travis-ci.org/symengine/symengine.R.svg?branch=master)](https://travis-ci.org/symengine/symengine.R)
[![AppVeyor Build status](https://ci.appveyor.com/api/projects/status/rr0tdh8ykvs04qg2?svg=true)](https://ci.appveyor.com/project/symengine/symengine-r)

This is an experiment to provide a R interface to the
[SymEngine library](https://github.com/symengine/symengine).

This project was a GSoC 2018 project under the organization
of The R Project for Statistical Computing.

## Installation

On Unix system, you may need install some system dependencies and compile the
symengine library manually. You can follow the instruction at 
https://github.com/symengine/symengine.

Alternatively, we have provided a convenient script to compile the symengine
library and install it to a non-root directory [`~/.local/rdeplibs-symengine`].
To use the script, first make sure you have installed `cmake`, `gmp` library,
`mpfr` library and `mpc` library on your system. For example, on openSUSE, Debian
or Mac OS, you will need:

```
zypper install cmake gmp-devel mpfr-devel mpc-devel    ## openSUSE
apt    install cmake libgmp-dev libmpfr-dev libmpc-dev ## Debian
brew   install cmake gmp mpfr libmpc                   ## Mac OS
```

Then in R, source the script (*Be aware of the risks running a script from online sources*)
and press Enter to select the default options:

```r
source("https://raw.githubusercontent.com/symengine/symengine.R/master/tools/install-symengine.R")
```

If it was successful, you can install the R package with

```{r, eval=FALSE}
devtools::install_github("symengine/symengine.R")
```

On Windows, the dependencies will be downloaded at build time, and you can install
directly with `devtools`.

Please report any problem installing the package on your system.

```{r}
library(symengine)
```

## Usage

### Manipulating Symbolic Expressions

```{r}
use_vars(x, y, z)
expr <- (x + y + z) ^ 2L - 42L
expand(expr)
```

Substitue `z` as `a` and `y` as `x^2`.

```{r}
expr <- subs(expr, ~ z, ~ a,
                   "y",   x^2L)
expr
```

Second derivative of `expr` with regards to `x`:

```{r}
d1_expr <- D(expr, "x")
d2_expr <- D(d1_expr, "x")
expand(d2_expr)
```

Solve the equation of `d2_expr == 0` with regards to `x`.

```{r}
solutions <- solve(d2_expr, "x")
solutions
```

### Numerically Evaluate Symbolic Expressions

For the two solutions above, we can convert them into a function that gives numeric
output with regards to given input.

```{r}
func <- as.function(solutions)
ans <- func(a = -100:-95)
colnames(ans) <- c("Solution1", "Solution2")
ans
```


### Numbers

The next prime number greater than 2^400.

```{r}
n <- nextprime(S(~ 2 ^ 400))
n
```

The greatest common divisor between the prime number and 42.

```{r}
GCD(n, 42)
```

The binomial coefficient `(2^30 ¦ 5)`.

```{r}
choose(S(~ 2^30), 5L)
```

Pi "computed" to 400-bit precision number.

```{r}
if (symengine_have_component("mpfr"))
    evalf(Constant("pi"), bits = 400)
```

### Object Equality

```{r}
x + y == S("x + y")
x + y != S("x + y")
```

```{r}
sin(x)/cos(x)
tan(x) == sin(x)/cos(x) # Different internal representation
```



## Related R Packages

- There are several functions in base R for defferentiation, integration, solving
  system of equations, etc. E.g. `solve`, `stats::D`, `stats::deriv`, `stats::integrate`,
  `stats::numericDeriv`.

- R package [`Deriv`](https://github.com/sgsokol/Deriv) for symbolic differentiation,
  it allows user to supply custom rules for differentiation.
- R package `numDeriv` for calculating numerical approximations to derivatives.

- R package `gmp` and `Rmpfr` provide multiple precision arithmetic and floating
  point operations. They also include some special functions, e.g. `Rmpfr::integrateR`
  for numerical integration.

- R package `mpc` available at [R forge](http://mpc.r-forge.r-project.org/). It
  provides multiple precision arithmetic for complex numbers.

- R package [`rSymPy`](https://cran.r-project.org/web/packages/rSymPy/index.html)
  provides an interface to 'SymPy' library in python via rJava.
- R package [`Ryacas`](https://cran.r-project.org/web/packages/Ryacas/index.html)
  provides an interface to the 'Yacas' computer algebra system.
  It is easier to install compared to `rSymPy`.


## Notes on some dependencies

The SymEngine library can optionally depend on some external libraries, which is
configured by CMake, see the list of CMake options in
[README of SymEngine](https://github.com/symengine/symengine/README.md) and
the
[configure script](https://github.com/Marlin-Na/Rlibsymengine/blob/master/configure)
of Rlibsymengine.

A few notes:

1. `GMP` (GNU Multiple Precision Arithmetic Library) is a C library that can be
    used to store and do arithmetic calculation with big integers and rationals.
    It has an R interface 
    ([gmp](https://github.com/cran/gmp/blob/master/DESCRIPTION) package).

2. `mpfr` (Multiple Precision Floating-Point Reliable) is a C library that depends
    on the `GMP` library and is used for arbitrary precision floating number
    arithmetic calculations.
    It has an R interface ([Rmpfr](https://github.com/cran/Rmpfr) package).
    This is an optional dependency for SymEngine.

3. `mpc` () is a C library that extends the `mpfr` library for the arithmetic of
    complex numbers with arbitrarily precision. There is a R package `mpc` which
    is not on CRAN, but [available](http://mpc.r-forge.r-project.org/) at R forge.

