---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# symengine


[![Travis-CI Build Status](https://travis-ci.org/Marlin-Na/symengine.R.svg?branch=master)](https://travis-ci.org/Marlin-Na/symengine.R)


This is an experiment to provide a R interface to the
[SymEngine library](https://github.com/symengine/symengine).
It is still in progress, but if you are interested, please contact
Jialin Ma <marlin-@gmx.cn> and Isuru Fernando <isuruf@gmail.com>.

This project is expected to be a GSoC 2018 project under the organization
of The R Project for Statistical Computing.


## Installation

Currently you need to manually compile and install
[SymEngine](https://github.com/symengine/symengine/) on your computer first
(Windows may not be supported yet).
For example, with the following steps:

First, you will need to install prerequisites and some optional dependencies.
For example on Debian-based system:

```sh
sudo apt-get install cmake libgmp-dev libmpfr-dev libmpc-dev binutils-dev
```

Second, download the source of symengine.

```sh
git clone git@github.com:symengine/symengine.git  # Download the source
cd symengine
```

Then use `cmake` to configure the compilation options and installation path.
SymEngine can be compiled with support of external libraries, such as mpfr,
mpc, etc. See the full list of available `cmake` options at
https://github.com/symengine/symengine/.
For example, use the following command:

```sh
cmake -DCMAKE_INSTALL_PREFIX:PATH="/usr/local" \ # Default installation prefix
    -DCMAKE_BUILD_TYPE=Debug \
    -DWITH_SYMENGINE_ASSERT=ON \
    -DWITH_MPFR=ON \
    -DWITH_MPC=ON \
    -DWITH_OPENMP=ON \
    -DWITH_BFD=yes \
    .
```

Then compile and install SymEngine to your computer with:

```
make
sudo make install
```

Once successful, install the R package from github with:

```{r, eval=FALSE}
devtools::install_github("Marlin-Na/symengine.R")
```

```{r}
library(symengine)
```

In the future, the SymEngine library may be contained in the package so that
there will be no need to manually install it.

If you run into problems installing the package, please let me know.

## Usage

### Symbol

A symbol or variable can be constructed from character.

```{r}
(x <- Symbol("x"))
(y <- Symbol("y"))
x ^ y
```

### Integer, RealDouble, RealMPFR

There are explicit constructors for such types:

```{r}
Integer(42L)
RealDouble(base::pi)
```

For large integer and high-precision floating number that R can not hold,
you can construct "Integer" or "RealMPFR" from character. For example:

```{r}
8615937169318
as.integer(8615937169318)
Integer("8615937169318")
Integer("8615937169318") ^ 9L
```

```{r, eval=FALSE}
# TODO, currently not available
RealMPFR("3.1415926535897932384626433832795028841971693993751058209", bits = 70)
```

Comparing with the `mpfr` function in `Rmpfr` package:

```{r}
Rmpfr::mpfr("3.1415926535897932384626433832795028841971693993751058209", precbits = 70)
```

Or simply use the SymEngine parser instead of the explicit constructors:

```{r}
S("8615937169318")
S("3.1415926535897932384626433832795028841971693993751058209")
```

### Complex, ComplexDouble and ComplexMPC

There will be explicit constructors (TODO):

```{r}
```

Or use the parser:

```{r}
S("6 + 9I")
```

Or use:

```{r}
6L + 9L * BuiltinConstant("I")
```

The `mpc` library is used for holding complex number with arbitrary precision,
similar to `mpfr` library for floating number.

```{r}
S("2.3 + 23.9999999999999999999I")
```

### Constants

For example:

```{r}
Constant("pi")
sin(Constant("pi") / 2L)
```

### Generic Conversion and Parser

As already showed in the above examples, `S` converts a R object to SymEngine object.
When the input is a character, it will parse the expression to produce appropriate object.

```{r}
S(6L)
S(4.2)
```

```{r}
(x <- S("x"))
(k <- S(~ k))          # Currently only work with "symbol"
(b <- Constant("b"))
S("k * x + b")
k * x + b
```

```{r}
S("pi")
S("sin(pi)")
```

```{r}
S("(tan(x) + sin(x))^2")
S("a + 2 >= a")
```

### Substitue a Variable

Substitute a variable with another one:

```{r}
x <- S("x")
a <- S("a")
(expr <- (tan(x) + sin(x) + a) ^ 2L)
```

```{r}
subs(expr, "x", "a")
subs(expr, "x", 3.1415926)
subs(expr, "x", Constant("pi"))
subs(expr, "x", Constant("pi") * 2L/3L)
```

### Expand an Expression

```{r}
expr
expand(expr)
```

### Derivative

```{r}
expr
diff(expr, "x")
diff(expr, "a")
```

### Evaluate an Expression

```{r}
evalf(Constant("pi"), bits = 999)
```

### Object Equality and Hash

```{r}
Eq(x + y, S("x + y"))
Neq(x + y, S("x + y"))
```

```{r}
tan(x)
sin(x)/cos(x)
Eq(tan(x), sin(x)/cos(x)) # Different internal representation
```

```{r}
Hash(tan(x))
Hash(sin(x)/cos(x))
```

### N Theory

TODO

### FunctionSymbol

TODO

```{r}
S("f(x, y)")
```

### Lambdify

TODO

### Vector

TODO

### Matrix (DenseMatrix and SparseMatrix)

TODO


## Under the Hood

The SymEngine objects are implemented with "externalptr":

```{r}
x <- Symbol(~ x)
str(x)
```

See "src/interface.c" for the C code that wraps the symengine api.


## Related R Packages

- There are several functions in base R for defferentiation, integration, solving
  system of equations, etc. E.g. `solve`, `stats::D`, `stats::deriv`, `stats::integrate`,
  `stats::numericDeriv`.

- R package [`Deriv`](https://github.com/sgsokol/Deriv) for symbolic differentiation,
  it allows user to supply custom rules for differentiation.
- R package `numDeriv` for calculating numerical approximations to derivatives.

- R package `gmp` and `Rmpfr` provide multiple precision arithmetic and floating
  point operations. They also include some special functions, e.g. `Rmpfr::integrateR`
  for numerical integration.

- R package `mpc` available at [R forge](http://mpc.r-forge.r-project.org/). It
  provides multiple precision arithmetic for complex numbers.

- R package [`rSymPy`](https://cran.r-project.org/web/packages/rSymPy/index.html)
  provides an interface to 'SymPy' library in python via rJava.
- R package [`Ryacas`](https://cran.r-project.org/web/packages/Ryacas/index.html)
  provides an interface to the 'Yacas' computer algebra system.
  It is easier to install compared to `rSymPy`.

