---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# symengine

[![Travis-CI Build Status](https://travis-ci.org/symengine/symengine.R.svg?branch=master)](https://travis-ci.org/symengine/symengine.R)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/Marlin-Na/symengine.R?branch=master&svg=true)](https://ci.appveyor.com/project/Marlin-Na/symengine.R)

This is an experiment to provide a R interface to the
[SymEngine library](https://github.com/symengine/symengine).
It is still in progress, but if you are interested, please contact
Jialin Ma <marlin-@gmx.cn> and Isuru Fernando <isuruf@gmail.com>.

This project was a GSoC 2018 project under the organization
of The R Project for Statistical Computing.

## Installation

Currently, you may need to install or compile symengine library manually on Linux or Mac OS.
Please follow the instruction at https://github.com/symengine/symengine.

Then in R, try

```{r, eval=FALSE}
devtools::install_github("symengine/symengine.R")
```

Please report any problem installing the package on your system.

```{r}
library(symengine)
```


## Usage

### Symbol

A symbol or variable can be constructed from character.

```{r}
(x <- Symbol("x"))
(y <- Symbol("y"))
x ^ y
```

### Integer, RealDouble, RealMPFR

There are explicit constructors for such types:

```{r}
Integer(42L)
RealDouble(base::pi)
```

For large integer and high-precision floating number that R can not hold,
you can construct "Integer" or "RealMPFR" from character. For example:

```{r}
8615937169318
as.integer(8615937169318)
Integer("8615937169318")
Integer("8615937169318") ^ 9L
```

```{r, eval=FALSE}
# TODO, currently not available
RealMPFR("3.1415926535897932384626433832795028841971693993751058209", bits = 70)
```

Comparing with the `mpfr` function in `Rmpfr` package:

```{r}
Rmpfr::mpfr("3.1415926535897932384626433832795028841971693993751058209", precbits = 70)
```

Or simply use the SymEngine parser instead of the explicit constructors:

```{r}
S("8615937169318")
S("3.1415926535897932384626433832795028841971693993751058209")
```

### Complex, ComplexDouble and ComplexMPC

There will be explicit constructors (TODO):

```{r}
```

Or use the parser:

```{r}
S("6 + 9I")
```

Or use:

```{r}
6L + 9L * Constant("I")
```

The `mpc` library is used for holding complex number with arbitrary precision,
similar to `mpfr` library for floating number.

```{r}
S("2.3 + 23.9999999999999999999I")
```

### Constants

For example:

```{r}
Constant("pi")
sin(Constant("pi") / 2L)
```

### Generic Conversion and Parser

As already showed in the above examples, `S` converts a R object to SymEngine object.
When the input is a character, it will parse the expression to produce appropriate object.

```{r}
S(6L)
S(4.2)
```

```{r}
(x <- S("x"))
(k <- S(~ k))          # Currently only work with "symbol"
(b <- Constant("b"))
S("k * x + b")
k * x + b
```

```{r}
S("pi")
S("sin(pi)")
```

```{r}
S("(tan(x) + sin(x))^2")
S("a + 2 >= a")
```

### Substitue a Variable

Substitute a variable with another one:

```{r}
x <- S("x")
a <- S("a")
(expr <- (tan(x) + sin(x) + a) ^ 2L)
```

```{r}
subs(expr, S("x"), S("a"))
subs(expr, S("x"), S(3.1415926))
subs(expr, S("x"), Constant("pi"))
subs(expr, S("x"), Constant("pi") * 2L/3L)
```

### Expand an Expression

```{r}
expr
expand(expr)
```

### Derivative

```{r}
expr
diff(expr, S("x"))
diff(expr, S("a"))
```

### Evaluate an Expression

```{r}
evalf(Constant("pi"), bits = 999)
```

### Object Equality and Hash

```{r}
x + y == S("x + y")
x + y != S("x + y")
```

```{r}
tan(x)
sin(x)/cos(x)
tan(x) == sin(x)/cos(x) # Different internal representation
```

```{r}
Hash(tan(x))
Hash(sin(x)/cos(x))
```

### N Theory

TODO

### FunctionSymbol

TODO

```{r}
S("f(x, y)")
```

### Lambdify

TODO

### Vector

```{r}
v <- vecbasic("a", 32L, 32)
v
v[-1]
v[[1]]
c(v, S("a + b"))
```

### Matrix (DenseMatrix and SparseMatrix)

TODO


## Under the Hood

The SymEngine objects are implemented with "externalptr":

```{r}
x <- Symbol(~ x)
str(x)
```

See "src/interface.c" for the C code that wraps the symengine api.


## Related R Packages

- There are several functions in base R for defferentiation, integration, solving
  system of equations, etc. E.g. `solve`, `stats::D`, `stats::deriv`, `stats::integrate`,
  `stats::numericDeriv`.

- R package [`Deriv`](https://github.com/sgsokol/Deriv) for symbolic differentiation,
  it allows user to supply custom rules for differentiation.
- R package `numDeriv` for calculating numerical approximations to derivatives.

- R package `gmp` and `Rmpfr` provide multiple precision arithmetic and floating
  point operations. They also include some special functions, e.g. `Rmpfr::integrateR`
  for numerical integration.

- R package `mpc` available at [R forge](http://mpc.r-forge.r-project.org/). It
  provides multiple precision arithmetic for complex numbers.

- R package [`rSymPy`](https://cran.r-project.org/web/packages/rSymPy/index.html)
  provides an interface to 'SymPy' library in python via rJava.
- R package [`Ryacas`](https://cran.r-project.org/web/packages/Ryacas/index.html)
  provides an interface to the 'Yacas' computer algebra system.
  It is easier to install compared to `rSymPy`.


## Notes on some dependencies

The SymEngine library can optionally depend on some external libraries, which is
configured by CMake, see the list of CMake options in
[README of SymEngine](https://github.com/symengine/symengine/README.md) and
the
[configure script](https://github.com/Marlin-Na/Rlibsymengine/blob/master/configure)
of Rlibsymengine.

A few notes:

1. `GMP` (GNU Multiple Precision Arithmetic Library) is a C library that can be
    used to store and do arithmetic calculation with big integers and rationals.
    It has an R interface 
    ([gmp](https://github.com/cran/gmp/blob/master/DESCRIPTION) package).

2. `mpfr` (Multiple Precision Floating-Point Reliable) is a C library that depends
    on the `GMP` library and is used for arbitrary precision floating number
    arithmetic calculations.
    It has an R interface ([Rmpfr](https://github.com/cran/Rmpfr) package).
    This is an optional dependency for SymEngine.

3. `mpc` () is a C library that extends the `mpfr` library for the arithmetic of
    complex numbers with arbitrarily precision. There is a R package `mpc` which
    is not on CRAN, but [available](http://mpc.r-forge.r-project.org/) at R forge.

