---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# symengine


[![Travis-CI Build Status](https://travis-ci.org/Marlin-Na/symengine.R.svg?branch=master)](https://travis-ci.org/Marlin-Na/symengine.R)


This is an experiment to provide a R interface to the
[SymEngine library](https://github.com/symengine/symengine).

Not fully functional yet, but if you are interested, please contact
Jialin Ma <marlin-@gmx.cn> and Isuru Fernando <isuruf@gmail.com>.

This project is expected to be a GSoC 2018 project under the organization
of The R Project for Statistical Computing.


## Installation

Currently you need to have SymEngine library installed on your computer, please
follow the instruction at https://github.com/symengine/symengine/, or do the following:

```sh
git clone git@github.com:symengine/symengine.git
cd symengine
make
sudo make install
```

In the future, the SymEngine library may be contained in the package so that no
external libraries will be needed.

Then install the R package from github with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Marlin-Na/symengine.R")
```

If you run into problems installing the package, please let me know.


## Example

Here are some examples showing the currently available functionalities.

```{r}
library(symengine)
S("k*x + b")     # Parse a string to expression
Integer(2^10)
Integer(2^10) ^ 9L
Integer(2^10) ^ 9L * Symbol("x")
x <- Symbol("x")
y <- Symbol("y")
4L * (x + y) - 2L * x
```

## TODO
